\documentclass[margin, 10pt]{res} % Use the res.cls style, the font size can be changed to 11pt or 12pt here

\usepackage{helvet} % Default font is the helvetica postscript font
%\usepackage{newcent} % To change the default font to the new century schoolbook postscript font uncomment this line and comment the one above

\setlength{\textwidth}{5.1in} % Text width of the document

\begin{document}

%----------------------------------------------------------------------------------------
%	NAME AND ADDRESS SECTION
%----------------------------------------------------------------------------------------

\moveleft.5\hoffset\centerline{\large\bf Clayton W. Seitz, Ph.D.} % Your name at the top
 
\moveleft\hoffset\vbox{\hrule width\resumewidth height 1pt}\smallskip % Horizontal line after name; adjust line thickness by changing the '1pt'
 
\moveleft.5\hoffset\centerline{cwseitz@iu.edu} % Your address
\moveleft.5\hoffset\centerline{cwseitz.github.io} % Your address

%----------------------------------------------------------------------------------------

\begin{resume}


\section{PERSONAL STATEMENT}  

%I have a broad background in machine learning, statistics, and physics/mathematics. Recently, I have researched variational diffusion models. I also specialize in optimization problems and machine learning methods for time series analysis and generative modeling.

%I have a background in statistics, machine learning, and mathematics. Recently, I have researched variational inference with diffusion models. I specialize in Bayesian modeling and the application of machine learning methods for time series analysis and generative modeling.

%I have a background in statistics, machine learning, and mathematics. Recently, I have researched prediction models based on variational diffusion. I generally specialize in the development and application of machine learning methods to make predictions from data.

I have a background in computer vision, machine learning, and mathematics. Recently, I have researched novel image generation methods based on deep learning. I also specialize in image/video processing, object detection, segmentation, and object tracking. These methods were applied in biological research.

%I have strong expertise in AI/ML techniques and frameworks, such as deep learning, generative modeling, and large language models. I also have proficiency in tools like Python, TensorFlow, PyTorch, and others. I am familiar cloud infrastructure, big data, and best practices in AI/ML and data science communities. I can lead complex projects and deliver successful outcomes.

%I have a background in machine learning and mathematics. Recently, I have researched novel image sampling methods based on deep learning. I also have a background with spiking neural networks and machine learning algorithms for image/video processing, object detection, segmentation, and object tracking

%I am a biophysicist with experience developing AI/ML and statistical methods to solve biological problems. This experience is interdisciplinary and lies at the interface of biology, physics, and AI/ML. 

%I specialize in machine learning methods for computer vision, image processing, and generative modeling. 

%I am a physicist with a background in image science, optics, and optical sensing. I specialize in the development and application of advanced modeling of image formation for image processing, computer vision, time series analysis, and generative modeling. 

%I specialize in the development and application of machine learning methods for computer vision, large language models, time series analysis, and generative modeling. I maintain a strong background in functional and object-oriented programming with experience writing production quality software.

%----------------------------------------------------------------------------------------
%	EDUCATION SECTION
%----------------------------------------------------------------------------------------

\section{EDUCATION}

\textbf{Doctor of Philosopy, Physics}\hfill 2024 \\
Purdue University\\
%Thesis: \textit{Advancing super-resolution microscopy for quantitative in-vivo imaging of chromatin nanodomains}

\textbf{Master of Science, Physics} \hfill 2021\\
University of Chicago\\
%Thesis: \textit{Stable cell assembly formation in excitatory-inhibitory neural networks}


\textbf{Bachelor of Science, Physics}, Magna Cum Laude \hfill 2019\\
Indiana University\\
Minor: Mathematics 

\textbf{Bachelor of Science, Informatics (Math Focus)}, Magna Cum Laude \hfill 2019\\
Indiana University\\


%----------------------------------------------------------------------------------------
%	PROFESSIONAL EXPERIENCE SECTION
%----------------------------------------------------------------------------------------
 
\section{EXPERIENCE}

\textbf{Graduate Researcher} \hfill 2021-2024 \\
Purdue University, Indianapolis, IN

\begin{itemize} \itemsep -2pt % Reduce space between items

\item Designed diffusion models and generative adversarial networks (GANs) and general computer vision techniques (object detection, segmentation, etc.) in PyTorch for modeling image datasets in super-resolution fluorescence microscopy

\item Used large datasets to train and optimize image synthesis models to generate high-quality microscopy images

%\item Developed analysis pipelines for high-throughput spatial biology assays such as multiplexed immunofluorescence and spatial transcriptomics

\item Applied deep neural networks for depth estimation and three dimensional reconstruction of objects in microscopy images


%\item Solved optimization problems in the context of image analysis e.g., particle localization, super-resolution, and inference of particle dynamics


\item Applied general probabilistic models for high-dimensional imaging datasets and associated Bayesian methods for statistical inference tasks

%\item Developed Hidden Markov models of photoswitching dynamics and particle motion


\end{itemize}

\textbf{Graduate Researcher} \hfill 2020-2021 \\
University of Chicago, Chicago, IL

\begin{itemize} \itemsep -2pt % Reduce space between items

\item Investigated fundamental learning mechanisms in recurrent neural networks (RNNs) using dynamical models, mean-field theory, and time-series analysis. 

\item Designed and ran Monte Carlo simulations of spiking neural networks 
 
\end{itemize}
 
\textbf{Research Assistant} \hfill 2018-2020\\
Purdue University, Indianapolis, IN
\begin{itemize} \itemsep -2pt

\item Developed a scientific package in Python for high-throughput object detection and tracking
\item Managed the package lifecycle and user training throughout the laboratory

\end{itemize}

%\textbf{Information Systems Intern} \hfill 2016\\
%Liberty Mutual Insurance
%\begin{itemize} \itemsep -2pt
%\item Wrote Python scripts to automate attestation testing of web facing servers
%\item Assisted in configuration of web application firewalls (WAFs)
%\end{itemize} 



%----------------------------------------------------------------------------------------
% AWARDS SECTION
%---------------------------------------------------------------------------------------- 

\section{AWARDS}

{\sl NIH Graduate Training Fellowship} \hfill 2020 \\
University of Chicago, Chicago, IL

{\sl Travel Award and Lightning Talk Invitation} \hfill 2019 \\
Physical Sciences in Oncology - Minneapolis, MN

{\sl Hudson and Holland Scholarship for Diversity and Inclusion} \hfill 2013-2017 \\
Indiana University, Bloomington, IN 

{\sl Founders Scholar} \hfill 2013-2017 \\
Indiana University, Bloomington, IN 

{\sl Cigital Scholarship} \hfill 2016-2017 \\
Indiana University, Bloomington, IN 

\section{PUBLICATIONS}

\textbf{Clayton Seitz}\textsuperscript{\textdagger}, Donghong Fu\textsuperscript{\textdagger}, Mengyuan Liu, Hailan Ma, and Jing Liu. \textit{BRD4 phosphorylation regulates the structure of chromatin nanodomains}.\\ Physical Review Letters (In Review). https://doi.org/10.1101/2024.09.03.611057. 2024

\textbf{Clayton Seitz} and Jing Liu. \textit{Uncertainty-aware localization microscopy by variational diffusion}. In Review. 2024

\textbf{Clayton Seitz} and Jing Liu. \textit{Quantum enhanced localization microscopy with a single photon avalanche diode array}. In Review. 2024

Maelle Locatelli\textsuperscript{\textdagger}, Josh Lawrimore\textsuperscript{\textdagger}, Hua Lin\textsuperscript{\textdagger}, Sarvath Sanaullah, \textbf{Clayton Seitz}, Dave Segall, Paul Kefer, Salvador Moreno Naike, Benton Lietz, Rebecca Anderson, Julia Holmes, Chongli Yuan, George Holzwarth, Bloom Kerry, Jing Liu, Keith D Bonin, Pierre-Alexandre Vidi. \textit{DNA damage reduces heterogeneity and coherence of chromatin motions}. PNAS 12 July 2022; 119 (29): 1-11
\\
\\
Mengdi Zhang, \textbf{Clayton Seitz}, Garrick Chang, Fadil Iqbal, Hua Lin, and Jing Liu \textit{A guide for single-particle chromatin tracking in live cell nuclei}. Cell Biology International 15 January 2022; 46 (5): 683-700
\\
\\
Wenting Wu, Farooq Syed, Edward Simpson, Chih-Chun Lee, Jing Liu, Garrick Chang, Chuanpeng Dong, \textbf{Clayton Seitz}, Decio L. Eizirik, Raghavendra G. Mirmira, Yunlong Liu, Carmella Evans-Molina; \textit{Impact of Proinflammatory Cytokines on Alternative Splicing Patterns in Human Islets}. Diabetes 25 October 2021; 71 (1): 116–127

\textbf{Clayton Seitz}, Hailan Ma, and Jing Liu. \textit{Cytokine-induced transcriptional memory is evident in the kinetics of transcriptional bursts}. Biophysical Society Annual Conference 2022


\textbf{Clayton Seitz}, Hua Lin, Keith Bonin, Pierre-Alexandre Vidi, and Jing Liu. \textit{Quantifying the spatiotemporal dynamics of dUTP labeled chromatin during the DNA damage response}. Biophysical Society Annual Conference 2020


\section{SOFTWARE \\ SKILLS} 

{\sl Programming Languages \& Software:} 
Linux, Bash, Python, R, PyTorch, C/C++, SQL, LaTeX, Git, Docker, SLURM, AWS\\

\end{resume}

\end{document}