\documentclass[12pt]{article}
\usepackage{amsmath} % AMS Math Package
\usepackage{bm}
\usepackage{amsthm} % Theorem Formatting
\usepackage{amssymb}    % Math symbols such as \mathbb
\usepackage{graphicx} % Allows for eps images
\usepackage[dvips,letterpaper,margin=1in,bottom=0.7in]{geometry}
\usepackage{tensor}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{physics}
\usepackage{amsmath, amssymb, graphics, setspace}

\newcommand{\mathsym}[1]{{}}
\newcommand{\unicode}[1]{{}}

\newcounter{mathematicapage}

\newtheorem{p}{Problem}
\usepackage{cancel}
\newtheorem*{lem}{Lemma}
\theoremstyle{definition}
\newtheorem*{dfn}{Definition}
 \newenvironment{s}{%\small%
        \begin{trivlist} \item \textbf{Solution}. }{%
            \hspace*{\fill} $\blacksquare$\end{trivlist}}%


\begin{document}

 {\noindent\Huge\bf  \\[0.5\baselineskip] {\fontfamily{cmr}\selectfont  Homework 7}         }\\[2\baselineskip] % Title
{ {\bf \fontfamily{cmr}\selectfont Quantum Mechanics}\\ {\textit{\fontfamily{cmr}\selectfont     \today}}}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~    {\large \textsc{C Seitz}
\\[1.4\baselineskip] 

\begin{p}
5.1
\end{p}

\begin{s}

We are concerned here with the new ground state ket $\ket{0}$ and the new ground state energy shift $\Delta_{0}$ in the presence of perturbation $V=bx$. 

\begin{align*}
\ket{0} = \ket{0^{0}} + \sum_{j\neq 0}\ket{j^{0}}\frac{V_{j0}}{E_{0}^{0}-E_{j}^{0}} + ...
\end{align*}

\begin{align*}
\Delta_{0} = V_{00} + \sum_{j\neq 0}\frac{|V_{j0}|^{2}}{E_{0}^{0}-E_{j}^{0}} + ...
\end{align*}

\begin{align*}
V_{nk} = b\bra{i^{0}}x\ket{j^{0}} &= b\sqrt{\frac{\hbar}{2m\omega}}\left(\sqrt{j}\delta_{i,j-1} + \sqrt{j+1}\delta_{i,j+1}\right)
\end{align*}

The lowest nonvanishing order is then $V_{01}$. Therefore 

\begin{align*}
\Delta_{0} = -\frac{b^{2}\hbar}{2m\omega}\frac{1}{\hbar\omega} = -\frac{b^{2}}{2m\omega^{2}}
\end{align*}

To solve it exactly, notice that the potential is of the form

\begin{align*}
V_{1}(x) = ax^{2} + bx
\end{align*}

The new potential shifts to the left by $b/2$, has a new minimum at $-b/2a$. So it is really just the original problem, we just have to make a change of coordinates and shift the equilibrium point down by $-b/2m\omega^{2}$. Therefore,

\begin{align*}
\Delta = -\frac{b}{2a} = -\frac{b}{2m\omega^{2}}
\end{align*}

which is exactly what we got with perturbation theory.

\end{s}

\begin{p}
5.2
\end{p}

\begin{s}

In general, the first order shift in the energy levels $i$ is

\begin{align*}
\Delta_{i} = V_{ii} = \bra{i^{0}}V\ket{i^{0}}
\end{align*}

Furthermore, the perturbation Hamiltonian is 


\begin{align*}
V = \frac{V_{0}x}{L}
\end{align*}

So we just need to identify the matrix elements along the diagonal of this matrix:

\begin{align*}
V_{ii} = \bra{i^{0}}V\ket{i^{0}} &= \frac{V_{0}}{L}\bra{i^{0}}x\ket{i^{0}} \\
&= \frac{V_{0}}{L}\frac{2}{L}\int_{0}^{L} x \sin^{2}\left(\frac{n\pi x}{L}\right)dx\\
&= \frac{V_{0}}{L}
\end{align*}

\end{s}

\begin{p}
5.5
\end{p}

\begin{s}

Up to order $\lambda^{2}$, we have

\begin{align*}
\ket{i} &= \ket{i^{0}} + \lambda\sum_{j\neq i}\frac{V_{ij}}{E_{i}^{0}-E_{j}^{0}}\ket{j^{0}} \\
&+ \lambda^{2}\left(\sum_{j\neq i}\sum_{l\neq i}\frac{V_{jl}V_{li}\ket{j^{0}}}{(E_{i}^{0}-E_{j}^{0})(E_{i}^{0}-E_{l}^{0})}-\sum_{j\neq i}\frac{V_{ii}V_{ji}\ket{j^{0}}}{(E_{i}^{0}-E_{j}^{0})^{2}}) \right)
\end{align*}

Now recall that we chose the normalization $\bra{i^{0}}\ket{i} = 1$ (which can be seen from the above equation), so what we have to calculate is $|\bra{i^{0}}\ket{i}|^{2}/|\bra{i}\ket{i}|^{2}$

\begin{align*}
\bra{i}\ket{i} &= 1 + \lambda^{2}\left(\sum_{j\neq i}\frac{V_{ij}^{*}}{E_{i}^{0}-E_{j}^{0}}\bra{j^{0}}\right)\left(\sum_{j\neq i}\frac{V_{ij}}{E_{i}^{0}-E_{j}^{0}}\ket{j^{0}}\right)\\
&= 1 + \lambda^{2}\sum_{j\neq i}\frac{V_{ij}^{*}V_{ij}}{\left(E_{i}^{0}-E_{j}^{0}\right)^{2}} 
\end{align*} 

So the probability is just

\begin{align*}
|\bra{i^{0}}\ket{i}|^{2}/|\bra{i}\ket{i}|^{2} = \left(1 + 2\lambda^{2}\sum_{j\neq i}\frac{V_{ij}^{*}V_{ij}}{\left(E_{i}^{0}-E_{j}^{0}\right)^{2}} \right)^{-1} + \mathcal{O}(\lambda^{3})
\end{align*} 

\end{s}

\begin{p}
5.7
\end{p}

\begin{s}

We can write this Hamiltonian as 
\begin{align*}
H_{0} = H_{x} + H_{y}
\end{align*}

Given some state $\ket{n, m}$, where $H_{x}$ acts on $\ket{n}$ and $H_{y}$ acts on $\ket{m}$. 

\begin{align*}
H_{0}\ket{n, m} &= (H_{x}+H_{y})\ket{n,m}\\
&= (E_{x}^{n}+E_{y}^{m})\ket{n,m}\\
&= \hbar\omega\left(n+\frac{1}{2} + m + \frac{1}{2}\right)\\
&= \hbar\omega(n + m + 1)
\end{align*}

So the energies of the three lowest states are $\hbar\omega, 2\hbar\omega, 3\hbar\omega$. There is a degeneracy - two unique $\ket{n,m}$ have the same eigenvalue for the first excited state. For example $\ket{0,1}$ and $\ket{1,0}$ have the same energy. Now, we are given the perturbation

\begin{align*}
V = \delta m\omega^{2}xy
\end{align*}

and we need to find a general matrix representation in the unperturbed $\ket{n,m}$ basis. I will use a shorthand:

\begin{align*}
V_{ij} &= \delta m\omega^{2}\bra{i}xy\ket{j}\\
&= \delta m\omega^{2}\bra{i_{x}}x\ket{j_{x}}\bra{i_{y}}y\ket{j_{y}}\\
&= \delta\frac{\hbar\omega}{2}\left(\sqrt{j_{x}}\delta_{i_{x},j_{x}-1} + \sqrt{j_{x}+1}\delta_{i_{x},j_{x}+1}\right)\left(\sqrt{j_{y}}\delta_{i_{y},j_{y}-1} + \sqrt{j_{y}+1}\delta_{i_{y},j_{y}+1}\right)
\end{align*}

First, we consider the ground state. To first order,

\begin{align*}
\Delta_{i} &= V_{ii} = 0
\end{align*}


This can be seen from the above equation, where the matrix elements of $V_{ij}$ are always zero when $i=j$. For the first exicted state, there is a degeneracy, so we need to diagonalize $V$ in the $\ket{1,0}$, $\ket{0,1}$ subspace. The matrix is

\begin{align*}
V' &= \begin{pmatrix}
\bra{0,1}V\ket{0,1} & \bra{0,1}V\ket{1,0}\\
\bra{1,0}V\ket{0,1} & \bra{1,0}V\ket{1,0}
\end{pmatrix}\\
&= \delta\frac{\hbar\omega}{2}\begin{pmatrix}
0 & 1\\
1 & 0
\end{pmatrix}
\end{align*}

which is a familiar matrix, so we can immediately write

\begin{align*}
\Delta_{1}^{1} = \pm \delta\frac{\hbar\omega}{2}
\end{align*}

and the eigenvectors are $\frac{1}{\sqrt{2}}\left(\ket{0,1}\pm\ket{1,0}\right)$


\end{s}

\begin{p}
5.12a
\end{p}

\begin{s}

We need to find the eigenvectors of $V$ in the degenerate subspace. If they are all independent, then the three-fold degeneracy is lifted. I will suppress the quantum numbers $n$ and $m_{s}$ for brevity

\begin{align*}
V &= \begin{pmatrix}
\bra{1,-1}V\ket{1,-1} & \bra{1,-1}V\ket{1,0} & \bra{1,-1}V\ket{1,1}\\
\bra{1,0}V\ket{1,-1} & \bra{1,0}V\ket{1,0} & \bra{1,0}V\ket{1,1}\\
\bra{1,1}V\ket{1,-1} & \bra{1,1}V\ket{1,0}& \bra{1,1}V\ket{1,1}
\end{pmatrix}\\
\end{align*}

It is immediately obvious that the elements along the diagonal must be zero, by symmetry. Also, the matrix is Hermitian (which can be verified by inspecting the spherical harmonics $Y_{l}^{m}(\theta,\phi)$. Therefore, we just need to find $\bra{1,-1}V\ket{1,0}$, $\bra{1,-1}V\ket{1,1}$ and $\bra{1,0}V\ket{1,1}$. Write,

\begin{align*}
\bra{1,-1}x^{2}\ket{1,0} &= \int r^{2}\sin^{2}\theta\cos^{2}\phi Y_{1}^{-1}(\theta,\phi)Y_{1}^{0}(\theta,\phi)rdrd\theta d\phi\\
&\propto \int r^{2}\sin^{2}\theta\cos^{2}\phi \sin\theta e^{i\phi}\cos\theta rdrd\theta d\phi = 0
\end{align*}

due to the $\theta$ integral. It is similar for $\langle y^{2} \rangle$ - the integral over $\theta$ is again zero and $\bra{1,-1}V\ket{1,0} = 0$. Moving on,

\begin{align*}
\bra{1,0}x^{2}\ket{1,1} &= \int r^{2}\sin^{2}\theta\cos^{2}\phi Y_{1}^{0}(\theta,\phi)Y_{1}^{1}(\theta,\phi)rdrd\theta d\phi\\
&\propto \int r^{2}\sin^{2}\theta\cos^{2}\phi \sin\theta e^{-i\phi}\cos\theta rdrd\theta d\phi = 0
\end{align*}

So it is basically the same as before and $\bra{1,0}V\ket{1,1} = 0$. Finally, 

\begin{align*}
\bra{1,-1}x^{2}\ket{1,1} &= \int r^{2}\sin^{2}\theta\cos^{2}\phi Y_{1}^{-1}(\theta,\phi)Y_{1}^{1}(\theta,\phi)rdrd\theta d\phi\\
&\propto \int r^{3}\sin^{4}\theta\cos^{2}\phi drd\theta d\phi
\end{align*}

\begin{align*}
\bra{1,-1}y^{2}\ket{1,1} &= \int r^{2}\sin^{2}\theta\cos^{2}\phi Y_{1}^{-1}(\theta,\phi)Y_{1}^{1}(\theta,\phi)rdrd\theta d\phi\\
&\propto \int r^{3}\sin^{2}\theta\cos^{4}\phi drd\theta d\phi
\end{align*}

We are not asked to work this integral out in detail, but it is nonzero and we will ultimately get a matrix like

\begin{align*}
V &= \begin{pmatrix}
0 & 0 & \alpha\\
0 & 0 & 0\\
\alpha & 0& 0
\end{pmatrix}
\end{align*}

since $\alpha$ is real. Finally, 

\begin{align*}
\mathrm{det}(V-\lambda I) = \mathrm{det}\begin{pmatrix}
-\lambda & 0 & \alpha\\
0 & -\lambda & 0\\
\alpha & 0& -\lambda
\end{pmatrix} = -\lambda^{3} + \lambda\alpha^{2} = 0
\end{align*}

which has three solutions $\lambda = 0,\pm\alpha$. So the degeneracy is broken.

\end{s}

\begin{p}
5.24
\end{p}

\begin{s}
\begin{align*}
V = \frac{A}{2\hbar^{2}}\left(\bm{J}^{2}-\bm{L}^{2}-\bm{S}^{2}\right) + \frac{B}{\hbar}\left(J_z + S_z\right)
\end{align*} 

We have two states for $l=0$ of form $\ket{lsjm}$

\begin{align*}
\ket{0, \frac{1}{2}, \frac{1}{2}, \frac{1}{2}},
\ket{0, \frac{1}{2}, \frac{1}{2}, -\frac{1}{2}}
\end{align*} 

and six states for $l=1$ 

\begin{align*}
\ket{1, \frac{1}{2}, \frac{3}{2}, \frac{3}{2}},
\ket{1, \frac{1}{2}, \frac{3}{2}, -\frac{3}{2}},
\ket{1, \frac{1}{2}, \frac{1}{2}, \frac{1}{2}},
\ket{1, \frac{1}{2}, \frac{3}{2}, \frac{1}{2}},
\ket{1, \frac{1}{2}, \frac{1}{2}, -\frac{1}{2}},
\ket{1, \frac{1}{2}, \frac{3}{2}, -\frac{1}{2}}
\end{align*} 

For an arbitrary state $\ket{lsjm}$

\begin{align*}
\bra{lsjm}V\ket{lsjm} &= \bra{lsjm}\frac{A}{2\hbar^{2}}\left(\bm{J}^{2}-\bm{L}^{2}-\bm{S}^{2}\right) + \frac{B}{\hbar}\left(J_z + S_z\right)\ket{lsjm}\\
&= \frac{A}{2}\left(j(j+1)+l(l+1)+s(s+1)\right) + B(m+s)
\end{align*} 

Suppose we group together the $l=0$ states into a 2x2 matrix

\begin{align*}
V_{1} &= \begin{pmatrix}
\frac{3A}{4} + B& 0\\
0 & \frac{3A}{4}\\
\end{pmatrix}
\end{align*}

\end{s}

\end{document}